---
title: "Installing packages for 'renv'"
author: "Ruben"
date: "2024-06-12"
categories: [code, quarto, github]
image: https://rstudio.github.io/renv/reference/figures/logo.svg
image-height: 230px
comments:
  giscus:
    repo: Carluna/bReaus-giscus
---

[`renv`](https://rstudio.github.io/renv/articles/renv.html){target="_blank"} is a R package that allows to create environments, where the user can install and manage packages. This is needed when using a GitHub action.

To use `renv` it is necessary to install R packages again, even though they are already installed locally.
This enables to use GitHub actions to automatically render the Quarto-Project to deploy a GitHub page!

Packages can be installed normally with `install.packages('tidyverse')`. Afterwards you can check the status of `renv` with `renv::status()`. If you installed new packages you have to run `renv::snapshot()`, to add the installed packages to the lockfile.

This is just a simple example to show, that the installation of the package worked and GitHub action runs:
```{r}
library(tidyverse)
```

```{r}
x <- seq(1,100,1)
y <- cos(x)

df <- tibble(x, y) |>
  mutate(z = sin(x))

head(df)
```

```{r}
ggplot(df) +
  geom_line(
    mapping = aes(x = x, y = y), linetype = 'dashed' 
  ) +
  geom_line(
    mapping = aes(x = x, y = z)
  ) +
  theme_minimal()
```

